Constructive Preference Elicitation by Setwise Max-margin Learning

### 2016-01-28 paper/ijcai16.tex
paper: figure finally within textpdflatex ijcai16
paper: getting closer to 7 pages..

\documentclass{article}
\usepackage{ijcai16}
\usepackage{times}
%\usepackage{latexsym}
%\usepackage{graphicx}
\usepackage{graphbox}
\usepackage{hyperref}
\usepackage{multirow}

@@

We compare {\sc SetMargin} against the following competitive Bayesian baselines:
% and \cite{viappiani2010optimal}.

\begin{itemize}
\item The Bayesian approach from \cite{guo2010real}, selecting queries
  according to {\em restricted informed VOI} ({\sc riVOI}), a
  computationally efficient heuristic approximation of
  value-of-information, and inference using TrueSkill$^{TM}$
  \cite{HerbrichMG06} (an inference scheme based on expectation
  propagation \cite{Minka01}).

 \item The Bayesian framework of \cite{viappiani2010optimal} using Monte Carlo methods
 \footnote{We use 50000 particles to represent the probability distribution.}for Bayesian inference and asking choice queries (asking the user to select its preferred item among a set of diplayed ones); the next query is computed by performing a greedy optimization of {\em Expected Utility of a Selection} (hereafter just called {\sc EUS}), a submodular quantity that can be used as a proxy for value of information (giving strong approximate guarantees).

 \item {\em Query Iteration} (referred as {\sc QI} below), also from \cite{viappiani2010optimal}, an even faster query selection method based on sampling a set of utility vectors, selecting the items that maximize utility for each of the vector and forming a first proposal for a query set. Then, a new set is computed by considering the optimal items in the associated  beliefs, and the process is repeated until convergence.
\end{itemize}

@@

@@

% \begin{figure*}[h!]
%     \centering
%     {\footnotesize
%     \begin{tabular}{cccc}
%         % \hline
%         % {\sc synthetic 3 loss} & {\sc synthetic 3 time} & {\sc synthetic 4 loss} & {\sc synthetic 4 time}
%         % \\
%         % \hline \hline
%         \multicolumn{4}{c}{{\sc Uniform}}
%         \\
%         \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_per_iter_loss} &
%         \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_per_iter_time} &
%         \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_per_iter_loss} &
%         \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_per_iter_time}
%         \\
%         \hline
%         \multicolumn{4}{c}{{\sc Normal}}
%         \\
%         \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_per_iter_loss} &
%         \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_per_iter_time} &
%         \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_per_iter_loss} &
%         \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_per_iter_time}
%         \\
%         \hline
%         \multicolumn{4}{c}{{\sc Sparse Uniform}}
%         \\
%         \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_sparse_per_iter_loss} &
%         \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_sparse_per_iter_time} &
%         \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_sparse_per_iter_loss} &
%         \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_sparse_per_iter_time}
%         \\
%         \hline
%         \multicolumn{4}{c}{{\sc Sparse Normal}}
%         \\
%         \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_sparse_per_iter_loss} &
%         \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_sparse_per_iter_time} &
%         \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_sparse_per_iter_loss} &
%         \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_sparse_per_iter_time}
%         \\
%         \hline
%     \end{tabular}
%     }
%     \caption{\label{fig:comparison} Comparison between {\sc SetMargin}
%       (orange), {\sc riVOI} (blue), {\sc QI} (green) and {\sc EUS}
%       (gray) on the $r=3$ (left) and $r=4$ (right) datasets. Each row
%       represents a different sampling distribution for user utility.
%       The number of iterations is plotted against the utility loss
%       (first and third columns) and the cumulative time (second and
%       fourth columns). Thick lines indicate median values over users,
%       while standard deviations are shown as shaded areas.}
% \end{figure*}

\begin{figure*}[h!]
    \centering
    {\footnotesize
    \begin{tabular}{ccccc}
        % \hline
        % {\sc synthetic 3 loss} & {\sc synthetic 3 time} & {\sc synthetic 4 loss} & {\sc synthetic 4 time}
        % \\
        % \hline \hline
        % \multicolumn{4}{c}{{\sc Uniform}}
        \\
        {\sc Uniform}  &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_uniform_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_uniform_per_iter_time} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_uniform_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_uniform_per_iter_time}
        \\
        \hline
         % \multicolumn{4}{c}{{\sc Normal}}
        \\
         {\sc Normal} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_normal_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_normal_per_iter_time} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_normal_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_normal_per_iter_time}
        \\
        \hline
        % \multicolumn{4}{c}{{\sc Sparse Uniform}}
        \\
        {\sc Sparse Uniform} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_uniform_sparse_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_uniform_sparse_per_iter_time} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_uniform_sparse_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_uniform_sparse_per_iter_time}
        \\
        \hline
        % \multicolumn{4}{c}{{\sc Sparse Normal}}
        \\
        {\sc Sparse Normal} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_normal_sparse_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_normal_sparse_per_iter_time} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_normal_sparse_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_normal_sparse_per_iter_time}
        \\
        \hline
    \end{tabular}
    }
    \caption{\label{fig:comparison} Comparison between {\sc SetMargin}
      (orange), {\sc riVOI} (blue), {\sc QI} (green) and {\sc EUS}
      (gray) on the $r=3$ (left) and $r=4$ (right) datasets. Each row
      represents a different sampling distribution for user utility.
      The number of iterations is plotted against the utility loss
      (first and third columns) and the cumulative time (second and
      fourth columns). Thick lines indicate median values over users,
      while standard deviations are shown as shaded areas.}
\end{figure*}

%\twocolumn

In Figure~\ref{fig:comparison} we report solution quality and timing
values for increasing number of collected user responses, for the
different competitors on each of the four different utility vector
distributions and datasets $r=3$ and $r=4$. Solution quality is
measured in terms of utility loss:
%
$$ \max_{\vx\in\calX_\text{feasible}} \left( u(\vx) - u(\vx^*) \right) $$
%
where $u(\cdot)$ is the true unknown user utility, and $\vx^*$ is the
solution recommended to the user after the elicitation phase (see
Algorithm~\ref{alg:setmargin}). Computational cost is measured in
terms of cumulative time. Given that {\sc riVOI},  {\sc QI} and
{\sc EUS} are single-threaded, we disabled multi-threading when running our
algorithm in these comparisons. All experiments were run on a 2.8 GHz Intel
Xeon CPU with 8 cores and 32 GiB of RAM.
%In all figures, thick lines indicate median values over the different
%users, shaded areas indicate standard deviations. Our {\sc SetMargin}
%algorithm for $k=2$ is reported in orange, {\sc riVOI} in blue, {\sc QI}
%in green and {\sc EUS} in gray. 
For all algorithms, one iteration corresponds to a single pairwise (we
used {\sc SetMargin} with $k=2$).
%query to the user 
%Results for $r=3$ and $r=4$ are reported.  
For dense weight vector distributions (first two rows), our approach
achieves results which are indistinguishable from the competitors in a
fraction of their time.  Indeed, all Bayesian approaches\footnote{Note
  that for {\sc QI} and {\sc EUS}, the time spent for Bayesian
  inference greatly dominates the time spent for choosing the next
  query to ask.\andrea{@paolo: necessario?}} become quickly intractable for growing values of $r$,
%cannot scale over $r=4$ (for $r=5$, {\sc riVOI} did not
%finish a single user after 24 hours for, Viappiani QI takes 20 minutes
%per user), \andrea{20 minuti mi paiono pochi, sei sicuro?}  
while our algorithm can easily scale to much larger datasets, as will
be shown later on. For sparse weight vector distributions (last two
rows) our approach, in addition to being substantially faster on each
iteration, requires less queries in order to reach optimal
solutions. This is an expected result as the sparsification norm in
our formulation ($\| \vw \|_1$) is enforcing sparsity in the weights, 
while none of the other approaches is %explicitly 
designed to favour sparse functions.

@@

\begin{figure*}[h!]
    \centering
    {\footnotesize
    \begin{tabular}{ccccc}
        % \hline
        % \multicolumn{2}{c}{{\sc synthetic 4}} &
        % \multicolumn{2}{c}{{\sc synthetic 5}}
        % \\
        % {\sc loss per iter.} & {\sc loss per query} & {\sc loss per iter.} & {\sc loss per query}
        % \\
        % \hline \hline
        {\sc uniform} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_4_uniform_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_4_uniform_per_query_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_5_uniform_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_5_uniform_per_query_loss}
        \\
        \hline
        {\sc sparse normal} &        
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_4_normal_sparse_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_4_normal_sparse_per_query_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_5_normal_sparse_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_5_normal_sparse_per_query_loss}
        \\
        \hline
     \end{tabular}
    }
    \caption{\label{fig:selfcomparison} Comparison for {\sc SetMargin}
      with $k=2,3,4$, in orange, blue and green respectively for the
      $r=4$ (left) and $r=5$ (right) datasets using uniform (top row) and sparse normal (bottom row) distributions. Median and standard deviation utility loss values are reported for increasing number of iterations ($1^{st}$ and $3^{rd}$ columns) and pairwise queries ($2^{nd}$ and $4^{th}$ columns).}
\end{figure*}

@@

A PC configurations is defined by eight attributes: computer type (laptop,
desktop, or tower), manufacturer (8 choices), CPU model (37), monitor size (8), RAM amount (10), storage (10)
size, and price.
%See Table~\ref{tab:pcdataset} for an overivew of the attributes.
The price attribute is defined as a linear combination of the other
attributes: this is a fair modeling choice, as often the price of a PC is
well approximated by the sum of the price of its components plus a bias due to
branding.
%
Interactions between attributes are expressed as Horn clauses. An example
constraint might look like this: ``if the manufacturer is Apple, then the
CPU must be either a PowerPC G3 or a PowerPC G4''\footnote{Despite the components
included in the dataset, which are a bit outdated, the dataset itself is pretty
realistic.}, which can be encoded
into MILP form as
%
$ (1 - x_\text{Apple}) + x_\text{G3} + x_\text{G4} \ge 1 $.
%
Note that mutual exclusivity between all binary variables of the
one-hot encoding of an attribute ensures that only one of
$x_\text{G3}$ and $x_\text{G4}$ can be 1.  The dataset includes
constraints between the following attributes: manufacturer $\to$ type,
manufacturer $\to$ CPU, type $\to$ RAM amount, type $\to$ storage
size, type $\to$ monitor size, for 16 Horn constraints total. We do
not report the full list due to space limitations. Note that the
search space is of the order of hundreds of thousands of candidate
configurations, and is far beyond reach of existing Bayesian
approaches.

>>>>

\documentclass{article}
\usepackage{ijcai16}
\usepackage{times}
%\usepackage{latexsym}
%\usepackage{graphicx}
\usepackage{graphbox}
\usepackage{hyperref}
\usepackage{multirow}
\usepackage{placeins}

@@

We compare {\sc SetMargin} against three state-of-the-art Bayesian
approaches: i) the Bayesian approach from \cite{guo2010real},
selecting queries according to {\em restricted informed VOI} ({\sc
  riVOI}), a computationally efficient heuristic approximation of
value-of-information, and inference using TrueSkill$^{TM}$
\cite{HerbrichMG06} (an inference scheme based on expectation
propagation \cite{Minka01}); ii) the Bayesian framework of
\cite{viappiani2010optimal} using Monte Carlo methods \footnote{We use
  50000 particles to represent the probability distribution.}for
Bayesian inference and asking choice queries (asking the user to
select its preferred item among a set of diplayed ones); the next
query is computed by performing a greedy optimization of {\em Expected
  Utility of a Selection} (hereafter just called {\sc EUS}), a
submodular quantity that can be used as a proxy for value of
information (giving strong approximate guarantees); iii) {\em Query
  Iteration} (referred as {\sc QI} below), also from
\cite{viappiani2010optimal}, an even faster query selection method
based on sampling a set of utility vectors, selecting the items that
maximize utility for each of the vector and forming a first proposal
for a query set. Then, a new set is computed by considering the
optimal items in the associated beliefs, and the process is repeated
until convergence.

@@

\begin{figure*}[t]
    \centering
    {\footnotesize
    \begin{tabular}{ccccc}
        % \hline
        % {\sc synthetic 3 loss} & {\sc synthetic 3 time} & {\sc synthetic 4 loss} & {\sc synthetic 4 time}
        % \\
        % \hline \hline
        % \multicolumn{4}{c}{{\sc Uniform}}
        \\
        {\sc Uniform}  &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_uniform_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_uniform_per_iter_time} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_uniform_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_uniform_per_iter_time}
        \\
        \hline
         % \multicolumn{4}{c}{{\sc Normal}}
        \\
         {\sc Normal} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_normal_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_normal_per_iter_time} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_normal_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_normal_per_iter_time}
        \\
        \hline
        % \multicolumn{4}{c}{{\sc Sparse Uniform}}
        \\
        {\sc Sparse Uniform} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_uniform_sparse_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_uniform_sparse_per_iter_time} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_uniform_sparse_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_uniform_sparse_per_iter_time}
        \\
        \hline
        % \multicolumn{4}{c}{{\sc Sparse Normal}}
        \\
        {\sc Sparse Normal} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_normal_sparse_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_normal_sparse_per_iter_time} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_normal_sparse_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_normal_sparse_per_iter_time}
        \\
        \hline
    \end{tabular}
    }
    \caption{\label{fig:comparison} Comparison between {\sc SetMargin}
      (orange), {\sc riVOI} (blue), {\sc QI} (green) and {\sc EUS}
      (gray) on the $r=3$ (left) and $r=4$ (right) datasets. Each row
      represents a different sampling distribution for user utility.
      The number of iterations is plotted against the utility loss
      (first and third columns) and the cumulative time (second and
      fourth columns). Thick lines indicate median values over users,
      while standard deviations are shown as shaded areas.}
\end{figure*}

@@

In Figure~\ref{fig:comparison} we report solution quality and timing
values for increasing number of collected user responses, for the
different competitors on each of the four different utility vector
distributions and datasets $r=3$ and $r=4$. Solution quality is
measured in terms of utility loss $ \max_{\vx\in\calX_\text{feasible}} \left( u(\vx) - u(\vx^*) \right)$,
where $u(\cdot)$ is the true unknown user utility, and $\vx^*$ is the
solution recommended to the user after the elicitation phase (see
Algorithm~\ref{alg:setmargin}). Computational cost is measured in
terms of cumulative time. Given that {\sc riVOI},  {\sc QI} and
{\sc EUS} are single-threaded, we disabled multi-threading when running our
algorithm in these comparisons. All experiments were run on a 2.8 GHz Intel
Xeon CPU with 8 cores and 32 GiB of RAM.
%In all figures, thick lines indicate median values over the different
%users, shaded areas indicate standard deviations. Our {\sc SetMargin}
%algorithm for $k=2$ is reported in orange, {\sc riVOI} in blue, {\sc QI}
%in green and {\sc EUS} in gray. 
For all algorithms, one iteration corresponds to a single pairwise (we
used {\sc SetMargin} with $k=2$).
%query to the user 
%Results for $r=3$ and $r=4$ are reported.  
For dense weight vector distributions (first two rows), our approach
achieves results which are indistinguishable from the competitors in a
fraction of their time.  Indeed, all Bayesian approaches%
%\footnote{Note
%  that for {\sc QI} and {\sc EUS}, the time spent for Bayesian
%  inference greatly dominates the time spent for choosing the next
%  query to ask.\andrea{@paolo: necessario?}} 
 become quickly intractable for growing values of $r$,
%cannot scale over $r=4$ (for $r=5$, {\sc riVOI} did not
%finish a single user after 24 hours for, Viappiani QI takes 20 minutes
%per user), \andrea{20 minuti mi paiono pochi, sei sicuro?}  
while our algorithm can easily scale to much larger datasets, as will
be shown later on. For sparse weight vector distributions (last two
rows) our approach, in addition to being substantially faster on each
iteration, requires less queries in order to reach optimal
solutions. This is an expected result as the sparsification norm in
our formulation ($\| \vw \|_1$) is enforcing sparsity in the weights, 
while none of the other approaches is %explicitly 
designed to favour sparse functions.

@@

\begin{figure*}[t]
    \centering
    {\footnotesize
    \begin{tabular}{ccccc}
        % \hline
        % \multicolumn{2}{c}{{\sc synthetic 4}} &
        % \multicolumn{2}{c}{{\sc synthetic 5}}
        % \\
        % {\sc loss per iter.} & {\sc loss per query} & {\sc loss per iter.} & {\sc loss per query}
        % \\
        % \hline \hline
        {\sc uniform} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_4_uniform_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_4_uniform_per_query_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_5_uniform_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_5_uniform_per_query_loss}
        \\
        \hline
        {\sc sparse normal} &        
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_4_normal_sparse_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_4_normal_sparse_per_query_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_5_normal_sparse_per_iter_loss} &
        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_5_normal_sparse_per_query_loss}
        \\
        \hline
     \end{tabular}
    }
    \caption{\label{fig:selfcomparison} Comparison for {\sc SetMargin}
      with $k=2,3,4$, in orange, blue and green respectively for the
      $r=4$ (left) and $r=5$ (right) datasets using uniform (top row) and sparse normal (bottom row) distributions. Median and standard deviation utility loss values are reported for increasing number of iterations ($1^{st}$ and $3^{rd}$ columns) and pairwise queries ($2^{nd}$ and $4^{th}$ columns).}
\end{figure*}

@@

A PC configurations is defined by eight attributes: computer type (laptop,
desktop, or tower), manufacturer (8 choices), CPU model (37), monitor size (8), RAM amount (10), storage (10)
size, and price.
%See Table~\ref{tab:pcdataset} for an overivew of the attributes.
The price attribute is defined as a linear combination of the other
attributes: this is a fair modeling choice, as often the price of a PC is
well approximated by the sum of the price of its components plus a bias due to
branding.
%
Interactions between attributes are expressed as Horn clauses. An example
constraint might look like this: ``if the manufacturer is Apple, then the
CPU must be either a PowerPC G3 or a PowerPC G4''%
%\footnote{Despite the components
%included in the dataset, which are a bit outdated, the dataset itself is pretty
%realistic.} \andrea{necessario?}
, which can be encoded into MILP form as
%
$ (1 - x_\text{Apple}) + x_\text{G3} + x_\text{G4} \ge 1 $.
%
Note that mutual exclusivity between all binary variables of the
one-hot encoding of an attribute ensures that only one of
$x_\text{G3}$ and $x_\text{G4}$ can be 1.  The dataset includes
constraints between the following attributes: manufacturer $\to$ type,
manufacturer $\to$ CPU, type $\to$ RAM amount, type $\to$ storage
size, type $\to$ monitor size, for 16 Horn constraints total. We do
not report the full list due to space limitations. Note that the
search space is of the order of hundreds of thousands of candidate
configurations, and is far beyond reach of existing Bayesian
approaches.
